---
title: "Women's Soccer Data"
output: html_notebook
---

We'll use this notebook to explore some NCAA women's soccer data and [our questions](https://docs.google.com/spreadsheets/d/1kTSrnl1URmcexIWLDMWqDnp3q-c2OS45uifnhjp10n4/edit#gid=0). First, let's load the tidyverse library:

```{r}
library(tidyverse)
```

We'll want to start with 2022 match data, which is located in [this GitHub repository](https://github.com/dwillis/NCAAWomensSoccerData). We can load the match data directly from GitHub using `read_csv`:


```{r}
matches_2022 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2022.csv")
```

Let's take a look at this data to see what it contains, using `head` and `glimpse`:

```{r}
head(matches_2022)
glimpse(matches_2022)
```

Now, onto the questions using aggregates, which means `group_by` and `summarize`:

1. How many shots in total has Maryland attempted and how many have been on goal? Calculate a percentage of shots on goal.

```{r}
matches_2022 %>% 
  filter(team == "Maryland Terrapins, Terps") %>% 
  summarize(total_shots = sum(sh_att), total_quality_shots = sum(so_g)) %>% 
  mutate(percentage_quality_shots = total_quality_shots / total_shots * 100)
 
## Total Maryland shots attempted: 65; Total Maryland shots on goal: 32; Shot on goal percentage: 49.23% ##
 
```

2. What about the same shot stats from last season's first five games? You'll need to load the matches from 2021 - just change the season.

```{r}
matches_2021 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2021.csv")

matches_2021 %>% 
  filter(team == "Maryland Terrapins, Terps", date < "2021-09-05") %>% 
  group_by(team) %>% 
  summarize(total_shots = sum(sh_att), total_quality_shots = sum(so_g)) %>% 
  mutate(percentage_quality_shots = total_quality_shots / total_shots * 100)
 
## Last season, through the first five matches: Maryland total shots: 86; Maryland total quality shots: 34; Quality shot percentage: 39.53 % ##
```


3. What about draws - Maryland started with 4 ties. Have any other teams had as many?

```{r}
draws <- matches_2022 %>% 
  group_by(team) %>% 
  select(team, result) %>%
  filter(result == "T") %>% 
  summarize(total = n()) %>% 
  arrange(desc(total))

## Four teams -- NJIT, Pacific, St. Bonaventure and VCU -- all have 4 draws ##
```


4. How many total corner kicks has Maryland had so far and how many have its opponents have? Calculate the difference between the two totals for every team.

```{r}
corner_kicks <- matches_2022 %>% 
  select(team, corners, defensive_corners) %>% 
  filter(team == "Maryland Terrapins, Terps") %>% 
  summarize(md_corners = sum(corners), opp_corners = sum(defensive_corners))

## Maryland has had 31 corners, while its opponents have had 24 corners ##

all_corner_kicks <- matches_2022 %>% 
 select(team, corners, defensive_corners) %>% 
  group_by(team) %>%
  summarize(corners = sum(corners), defensive_corners = sum(defensive_corners)) %>% 
  mutate(corners_diff = corners - defensive_corners)

```
