---
title: "New Background"
output: html_notebook
---

## I changed my research question. I want to see which NFL team has the highest percentage of scoring on the ensuing drive after logging a takeaway. Essentially, I looked to log each team's total takeways (fumble recoveries, interceptions and turnovers on downs) from the 2002-2021 seasons and divide it by how many ensuing offensive drives ended with either a touchdown or a field goal. I could have done this more efficiently, but I created separate dataframes for all the fumbles, interceptions and turnovers on downs for each season from 2002 - 2021, grouped them by team and counted them. I created a dataframe called "teamtakeawayssince02" that totaled up each team's fumble recoveries, interceptions and fourth down stops thoughout the 20-season window. Then I used the NFLFastR dataframes to filter out touchdowns and field goals when the previous drive ended in either a fumble, interception or turnover on downs, combining the dataframes -- with each one representing a season -- into one big frame called 'scoresofftovs.' After that, I combined the two dataframes and mutated a column dividing the number of scores off takeaways by total takeaways. ##

```{r}
library(tidyverse)
library(nflfastR)
library(stringr)
```

```{r}
nfl21 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2021.rds")
```

```{r}
nfl21 %>% 
  group_by(fourth_down_failed) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
ints21 <- nfl21 %>% 
  select(week, posteam, defteam, drive, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "INTERCEPTION", str_detect(desc, "INTERCEPTED"))
```

```{r}
teamints21 <- ints21 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
nfl20 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2020.rds")
```

```{r}
ints20 <- nfl20 %>% 
  select(week, posteam, defteam, drive, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "INTERCEPTION", str_detect(desc, "INTERCEPTED"))
```

```{r}
teamints20 <- ints20 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
nfl19 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2019.rds")
```

```{r}
ints19 <- nfl19 %>% 
  select(week, posteam, defteam, drive, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "INTERCEPTION", str_detect(desc, "INTERCEPTED"))
```

```{r}
teamints19 <- ints19 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
nfl18 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2018.rds")
```

```{r}
ints18 <- nfl18 %>% 
  select(week, posteam, defteam, drive, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "INTERCEPTION", str_detect(desc, "INTERCEPTED"))
```

```{r}
teamints18 <- ints18 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
nfl17 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2017.rds")
```

```{r}
ints17 <- nfl17 %>% 
  select(week, posteam, defteam, drive, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "INTERCEPTION", str_detect(desc, "INTERCEPTED"))
```

```{r}
teamints17 <- ints17 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
nfl16 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2016.rds")
```

```{r}
ints16 <- nfl16 %>% 
  select(week, posteam, defteam, drive, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "INTERCEPTION", str_detect(desc, "INTERCEPTED"))
```

```{r}
teamints16 <- ints16 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
nfl15 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2015.rds")
```

```{r}
ints15 <- nfl15 %>% 
  select(week, posteam, defteam, drive, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "INTERCEPTION", str_detect(desc, "INTERCEPTED"))
```

```{r}
teamints15 <- ints15 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
nfl14 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2014.rds")
```

```{r}
ints14 <- nfl14 %>% 
  select(week, posteam, defteam, drive, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "INTERCEPTION", str_detect(desc, "INTERCEPTED"))
```

```{r}
teamints14 <- ints14 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
nfl13 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2013.rds")
```

```{r}
ints13 <- nfl13 %>% 
  select(week, posteam, defteam, drive, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "INTERCEPTION", str_detect(desc, "INTERCEPTED"))
```

```{r}
teamints13 <- ints13 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
nfl12 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2012.rds")
```

```{r}
ints12 <- nfl12 %>% 
  select(week, posteam, defteam, drive, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "INTERCEPTION", str_detect(desc, "INTERCEPTED"))
```

```{r}
teamints12 <- ints12 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
nfl11 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2011.rds")
```

```{r}
ints11 <- nfl11 %>% 
  select(week, posteam, defteam, drive, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "INTERCEPTION", str_detect(desc, "INTERCEPTED"))
```

```{r}
teamints11 <- ints11 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
nfl10 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2010.rds")
```

```{r}
ints10 <- nfl10 %>% 
  select(week, posteam, defteam, drive, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "INTERCEPTION", str_detect(desc, "INTERCEPTED"))
```

```{r}
teamints10 <- ints10 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
nfl09 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2009.rds")
```

```{r}
ints09 <- nfl09 %>% 
  select(week, posteam, defteam, drive, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "INTERCEPTION", str_detect(desc, "INTERCEPTED"))
```

```{r}
teamints09 <- ints09 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
nfl08 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2008.rds")
```

```{r}
ints08 <- nfl08 %>% 
  select(week, posteam, defteam, drive, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "INTERCEPTION", str_detect(desc, "INTERCEPTED"))
```

```{r}
teamints08 <- ints08 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
nfl07 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2007.rds")
```

```{r}
ints07 <- nfl07 %>% 
  select(week, posteam, defteam, drive, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "INTERCEPTION", str_detect(desc, "INTERCEPTED"))
```

```{r}
teamints07 <- ints07 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
nfl06 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2006.rds")
```

```{r}
ints06 <- nfl06 %>% 
  select(week, posteam, defteam, drive, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "INTERCEPTION", str_detect(desc, "INTERCEPTED"))
```

```{r}
teamints06 <- ints06 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
nfl05 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2005.rds")
```

```{r}
ints05 <- nfl05 %>% 
  select(week, posteam, defteam, drive, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "INTERCEPTION", str_detect(desc, "INTERCEPTED"))
```

```{r}
teamints05 <- ints05 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
nfl04 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2004.rds")
```

```{r}
ints04 <- nfl04 %>% 
  select(week, posteam, defteam, drive, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "INTERCEPTION", str_detect(desc, "INTERCEPTED"))
```

```{r}
teamints04 <- ints04 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
nfl03 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2003.rds")
```

```{r}
ints03 <- nfl03 %>% 
  select(week, posteam, defteam, drive, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "INTERCEPTION", str_detect(desc, "INTERCEPTED"))
```

```{r}
teamints03 <- ints03 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
nfl02 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2002.rds")
```

```{r}
ints02 <- nfl02 %>% 
  select(week, posteam, defteam, drive, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "INTERCEPTION", str_detect(desc, "INTERCEPTED"))
```

```{r}
teamints02 <- ints02 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
team_ints_since_02 <- rbind(teamints02, teamints03, teamints04, teamints05, teamints06, teamints07, teamints08, teamints09, teamints10, teamints11, teamints12, teamints13, teamints14, teamints15, teamints16, teamints17, teamints18, teamints19, teamints20, teamints21) %>% 
  group_by(defteam) %>%
  summarize(sum(total))
```

```{r}
fumbles21 <- nfl21 %>% 
    select(week, posteam, defteam, drive, series_result, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "FUMBLE", str_detect(desc, "RECOVERED"))
```

```{r}
teamfumbles21 <- fumbles21 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fumbles20 <- nfl20 %>% 
    select(week, posteam, defteam, drive, series_result, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "FUMBLE", str_detect(desc, "RECOVERED"))
```

```{r}
teamfumbles20 <- fumbles20 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fumbles19 <- nfl19 %>% 
    select(week, posteam, defteam, drive, series_result, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "FUMBLE", str_detect(desc, "RECOVERED"))
```

```{r}
teamfumbles19 <- fumbles19 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fumbles18 <- nfl18 %>% 
    select(week, posteam, defteam, drive, series_result, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "FUMBLE", str_detect(desc, "RECOVERED"))
```

```{r}
teamfumbles18 <- fumbles18 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fumbles17 <- nfl17 %>% 
    select(week, posteam, defteam, drive, series_result, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "FUMBLE", str_detect(desc, "RECOVERED"))
```

```{r}
teamfumbles17 <- fumbles17 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fumbles16 <- nfl16 %>% 
    select(week, posteam, defteam, drive, series_result, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "FUMBLE", str_detect(desc, "RECOVERED"))
```

```{r}
teamfumbles16 <- fumbles16 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fumbles15 <- nfl15 %>% 
    select(week, posteam, defteam, drive, series_result, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "FUMBLE", str_detect(desc, "RECOVERED"))
```

```{r}
teamfumbles15 <- fumbles15 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fumbles14 <- nfl14 %>% 
    select(week, posteam, defteam, drive, series_result, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "FUMBLE", str_detect(desc, "RECOVERED"))
```

```{r}
teamfumbles14 <- fumbles14 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fumbles13 <- nfl13 %>% 
    select(week, posteam, defteam, drive, series_result, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "FUMBLE", str_detect(desc, "RECOVERED"))
```

```{r}
teamfumbles13 <- fumbles13 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fumbles12 <- nfl12 %>% 
    select(week, posteam, defteam, drive, series_result, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "FUMBLE", str_detect(desc, "RECOVERED"))
```

```{r}
teamfumbles12 <- fumbles12 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fumbles11 <- nfl11 %>% 
    select(week, posteam, defteam, drive, series_result, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "FUMBLE", str_detect(desc, "RECOVERED"))
```

```{r}
teamfumbles11 <- fumbles11 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fumbles10 <- nfl10 %>% 
    select(week, posteam, defteam, drive, series_result, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "FUMBLE", str_detect(desc, "RECOVERED"))
```

```{r}
teamfumbles10 <- fumbles10 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fumbles09 <- nfl09 %>% 
    select(week, posteam, defteam, drive, series_result, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "FUMBLE", str_detect(desc, "RECOVERED"))
```

```{r}
teamfumbles09 <- fumbles09 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fumbles08 <- nfl08 %>% 
    select(week, posteam, defteam, drive, series_result, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "FUMBLE", str_detect(desc, "RECOVERED"))
```

```{r}
teamfumbles08 <- fumbles08 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fumbles07 <- nfl07 %>% 
    select(week, posteam, defteam, drive, series_result, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "FUMBLE", str_detect(desc, "RECOVERED"))
```

```{r}
teamfumbles07 <- fumbles07 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fumbles06 <- nfl06 %>% 
    select(week, posteam, defteam, drive, series_result, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "FUMBLE", str_detect(desc, "RECOVERED"))
```

```{r}
teamfumbles06 <- fumbles06 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fumbles05 <- nfl05 %>% 
    select(week, posteam, defteam, drive, series_result, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "FUMBLE", str_detect(desc, "RECOVERED"))
```

```{r}
teamfumbles05 <- fumbles05 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fumbles04 <- nfl04 %>% 
    select(week, posteam, defteam, drive, series_result, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "FUMBLE", str_detect(desc, "RECOVERED"))
```

```{r}
teamfumbles04 <- fumbles04 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fumbles03 <- nfl03 %>% 
    select(week, posteam, defteam, drive, series_result, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "FUMBLE", str_detect(desc, "RECOVERED"))
```

```{r}
teamfumbles03 <- fumbles03 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fumbles02 <- nfl02 %>% 
    select(week, posteam, defteam, drive, series_result, drive_start_transition, drive_end_transition, score_differential, desc) %>% 
  filter(drive_end_transition == "FUMBLE", str_detect(desc, "RECOVERED"))
```

```{r}
teamfumbles02 <- fumbles02 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
team_frs_since_02 <- rbind(teamfumbles02, teamfumbles03, teamfumbles04, teamfumbles05, teamfumbles06, teamfumbles07, teamfumbles08, teamfumbles09, teamfumbles10, teamfumbles11, teamfumbles12, teamfumbles13, teamfumbles14, teamfumbles15, teamfumbles16, teamfumbles17, teamfumbles18, teamfumbles19, teamfumbles20, teamfumbles21) %>%   
  group_by(defteam) %>%
  summarize(sum(total))
```

```{r}
fourthdownstops21 <- nfl21 %>% 
  select(defteam, week, posteam, drive, series_result, fourth_down_failed) %>% 
  filter(fourth_down_failed == 1) %>% 
  filter(series_result %in% "Turnover on downs")
```

```{r}
teamdowns21 <- fourthdownstops21 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fourthdownstops20 <- nfl20 %>% 
  select(defteam, week, posteam, drive, series_result, fourth_down_failed) %>% 
  filter(fourth_down_failed == 1) %>% 
  filter(series_result %in% "Turnover on downs")
```

```{r}
teamdowns20 <- fourthdownstops20 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fourthdownstops19 <- nfl19 %>% 
  select(defteam, week, posteam, drive, series_result, fourth_down_failed) %>% 
  filter(fourth_down_failed == 1) %>% 
  filter(series_result %in% "Turnover on downs")
```

```{r}
teamdowns19 <- fourthdownstops19 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fourthdownstops18 <- nfl18 %>% 
  select(defteam, week, posteam, drive, series_result, fourth_down_failed) %>% 
  filter(fourth_down_failed == 1) %>% 
  filter(series_result %in% "Turnover on downs")
```

```{r}
teamdowns18 <- fourthdownstops18 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fourthdownstops17 <- nfl17 %>% 
  select(defteam, week, posteam, drive, series_result, fourth_down_failed) %>% 
  filter(fourth_down_failed == 1) %>% 
  filter(series_result %in% "Turnover on downs")
```

```{r}
teamdowns17 <- fourthdownstops17 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fourthdownstops16 <- nfl16 %>% 
  select(defteam, week, posteam, drive, series_result, fourth_down_failed) %>% 
  filter(fourth_down_failed == 1) %>% 
  filter(series_result %in% "Turnover on downs")
```

```{r}
teamdowns16 <- fourthdownstops16 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fourthdownstops15 <- nfl15 %>% 
  select(defteam, week, posteam, drive, series_result, fourth_down_failed) %>% 
  filter(fourth_down_failed == 1) %>% 
  filter(series_result %in% "Turnover on downs")
```

```{r}
teamdowns15 <- fourthdownstops15 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fourthdownstops14 <- nfl14 %>% 
  select(defteam, week, posteam, drive, series_result, fourth_down_failed) %>% 
  filter(fourth_down_failed == 1) %>% 
  filter(series_result %in% "Turnover on downs")
```

```{r}
teamdowns14 <- fourthdownstops14 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fourthdownstops13 <- nfl13 %>% 
  select(defteam, week, posteam, drive, series_result, fourth_down_failed) %>% 
  filter(fourth_down_failed == 1) %>% 
  filter(series_result %in% "Turnover on downs")
```

```{r}
teamdowns13 <- fourthdownstops13 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fourthdownstops12 <- nfl12 %>% 
  select(defteam, week, posteam, drive, series_result, fourth_down_failed) %>% 
  filter(fourth_down_failed == 1) %>% 
  filter(series_result %in% "Turnover on downs")
```

```{r}
teamdowns12 <- fourthdownstops12 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fourthdownstops11 <- nfl11 %>% 
  select(defteam, week, posteam, drive, series_result, fourth_down_failed) %>% 
  filter(fourth_down_failed == 1) %>% 
  filter(series_result %in% "Turnover on downs")
```

```{r}
teamdowns11 <- fourthdownstops11 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fourthdownstops10 <- nfl10 %>% 
  select(defteam, week, posteam, drive, series_result, fourth_down_failed) %>% 
  filter(fourth_down_failed == 1) %>% 
  filter(series_result %in% "Turnover on downs")
```

```{r}
teamdowns10 <- fourthdownstops10 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fourthdownstops09 <- nfl09 %>% 
  select(defteam, week, posteam, drive, series_result, fourth_down_failed) %>% 
  filter(fourth_down_failed == 1) %>% 
  filter(series_result %in% "Turnover on downs")
```

```{r}
teamdowns09 <- fourthdownstops09 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fourthdownstops08 <- nfl08 %>% 
  select(defteam, week, posteam, drive, series_result, fourth_down_failed) %>% 
  filter(fourth_down_failed == 1) %>% 
  filter(series_result %in% "Turnover on downs")
```

```{r}
teamdowns08 <- fourthdownstops08 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fourthdownstops07 <- nfl07 %>% 
  select(defteam, week, posteam, drive, series_result, fourth_down_failed) %>% 
  filter(fourth_down_failed == 1) %>% 
  filter(series_result %in% "Turnover on downs")
```

```{r}
teamdowns07 <- fourthdownstops07 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fourthdownstops06 <- nfl06 %>% 
  select(defteam, week, posteam, drive, series_result, fourth_down_failed) %>% 
  filter(fourth_down_failed == 1) %>% 
  filter(series_result %in% "Turnover on downs")
```

```{r}
teamdowns06 <- fourthdownstops06 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fourthdownstops05 <- nfl05 %>% 
  select(defteam, week, posteam, drive, series_result, fourth_down_failed) %>% 
  filter(fourth_down_failed == 1) %>% 
  filter(series_result %in% "Turnover on downs")
```

```{r}
teamdowns05 <- fourthdownstops05 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fourthdownstops04 <- nfl04 %>% 
  select(defteam, week, posteam, drive, series_result, fourth_down_failed) %>% 
  filter(fourth_down_failed == 1) %>% 
  filter(series_result %in% "Turnover on downs")
```

```{r}
teamdowns04 <- fourthdownstops04 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fourthdownstops03 <- nfl03 %>% 
  select(defteam, week, posteam, drive, series_result, fourth_down_failed) %>% 
  filter(fourth_down_failed == 1) %>% 
  filter(series_result %in% "Turnover on downs")
```

```{r}
teamdowns03 <- fourthdownstops03 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fourthdownstops02 <- nfl02 %>% 
  select(defteam, week, posteam, drive, series_result, fourth_down_failed) %>% 
  filter(fourth_down_failed == 1) %>% 
  filter(series_result %in% "Turnover on downs")
```

```{r}
teamdowns02 <- fourthdownstops02 %>% 
  group_by(defteam) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total))
```

```{r}
fourth_down_stops_since_02 <- rbind(teamdowns02, teamdowns03, teamdowns04, teamdowns05, teamdowns06, teamdowns07, teamdowns08, teamdowns09, teamdowns10, teamdowns11, teamdowns12, teamdowns13, teamdowns14, teamdowns15, teamdowns16, teamdowns17, teamdowns18, teamdowns19, teamdowns20, teamdowns21) %>% 
    group_by(defteam) %>%
  summarize(sum(total))
```

```{r}
teamtakeawaysince02 <- rbind(fourth_down_stops_since_02, team_frs_since_02, team_ints_since_02) %>%  rename(takeaways='sum(total)') %>% 
group_by(defteam) %>% 
  summarize(sum(takeaways))
```

```{r}
nfl21 %>% 
filter(drive_start_transition %in% (tov), drive_ended_with_score == 1) %>% 
  filter(str_detect(desc, 'score')
```

```{r}
score <- c("field goal is GOOD", "TOUCHDOWN")
```

```{r}
scoresofftovs21 <- nfl21 %>% 
  filter(str_detect(desc, "is GOOD")) %>% 
  filter(drive_start_transition %in% c("INTERCEPTION", "FUMBLE", "DOWNS")) %>%
  filter(series_result %in% c("Touchdown", "Field goal")) %>% 
  select(posteam, defteam, drive_start_transition, desc, series_result) %>% 
  group_by(posteam) %>% 
  summarize(total=n())
```

```{r}
scoresofftovs20 <- nfl20 %>% 
  filter(str_detect(desc, "is GOOD")) %>% 
  filter(drive_start_transition %in% c("INTERCEPTION", "FUMBLE", "DOWNS")) %>%
  filter(series_result %in% c("Touchdown", "Field goal")) %>% 
  select(posteam, defteam, drive_start_transition, desc, series_result) %>% 
  group_by(posteam) %>% 
  summarize(total=n())
  
```

```{r}
scoresofftovs19 <- nfl19 %>% 
  filter(str_detect(desc, "is GOOD")) %>% 
  filter(drive_start_transition %in% c("INTERCEPTION", "FUMBLE", "DOWNS")) %>%
  filter(series_result %in% c("Touchdown", "Field goal")) %>% 
  select(posteam, defteam, drive_start_transition, desc, series_result) %>% 
  group_by(posteam) %>% 
  summarize(total=n())
```

```{r}
scoresofftovs18 <- nfl18 %>% 
  filter(str_detect(desc, "is GOOD")) %>% 
  filter(drive_start_transition %in% c("INTERCEPTION", "FUMBLE", "DOWNS")) %>%
  filter(series_result %in% c("Touchdown", "Field goal")) %>% 
  select(posteam, defteam, drive_start_transition, desc, series_result) %>% 
  group_by(posteam) %>% 
  summarize(total=n())
```

```{r}
scoresofftovs17 <- nfl17 %>% 
  filter(str_detect(desc, "is GOOD")) %>% 
  filter(drive_start_transition %in% c("INTERCEPTION", "FUMBLE", "DOWNS")) %>%
  filter(series_result %in% c("Touchdown", "Field goal")) %>% 
  select(posteam, defteam, drive_start_transition, desc, series_result) %>% 
  group_by(posteam) %>% 
  summarize(total=n())
```

```{r}
scoresofftovs16 <- nfl16 %>% 
  filter(str_detect(desc, "is GOOD")) %>% 
  filter(drive_start_transition %in% c("INTERCEPTION", "FUMBLE", "DOWNS")) %>%
  filter(series_result %in% c("Touchdown", "Field goal")) %>% 
  select(posteam, defteam, drive_start_transition, desc, series_result) %>% 
  group_by(posteam) %>% 
  summarize(total=n())
```

```{r}
scoresofftovs15 <- nfl15 %>% 
  filter(str_detect(desc, "is GOOD")) %>% 
  filter(drive_start_transition %in% c("INTERCEPTION", "FUMBLE", "DOWNS")) %>%
  filter(series_result %in% c("Touchdown", "Field goal")) %>% 
  select(posteam, defteam, drive_start_transition, desc, series_result) %>% 
  group_by(posteam) %>% 
  summarize(total=n())
```

```{r}
scoresofftovs14 <- nfl14 %>% 
  filter(str_detect(desc, "is GOOD")) %>% 
  filter(drive_start_transition %in% c("INTERCEPTION", "FUMBLE", "DOWNS")) %>%
  filter(series_result %in% c("Touchdown", "Field goal")) %>% 
  select(posteam, defteam, drive_start_transition, desc, series_result) %>% 
  group_by(posteam) %>% 
  summarize(total=n())
```

```{r}
scoresofftovs13 <- nfl13 %>% 
  filter(str_detect(desc, "is GOOD")) %>% 
  filter(drive_start_transition %in% c("INTERCEPTION", "FUMBLE", "DOWNS")) %>%
  filter(series_result %in% c("Touchdown", "Field goal")) %>% 
  select(posteam, defteam, drive_start_transition, desc, series_result) %>% 
  group_by(posteam) %>% 
  summarize(total=n())
```

```{r}
scoresofftovs12 <- nfl12 %>% 
  filter(str_detect(desc, "is GOOD")) %>% 
  filter(drive_start_transition %in% c("INTERCEPTION", "FUMBLE", "DOWNS")) %>%
  filter(series_result %in% c("Touchdown", "Field goal")) %>% 
  select(posteam, defteam, drive_start_transition, desc, series_result) %>% 
  group_by(posteam) %>% 
  summarize(total=n())
```

```{r}
scoresofftovs11 <- nfl11 %>% 
  filter(str_detect(desc, "is GOOD")) %>% 
  filter(drive_start_transition %in% c("INTERCEPTION", "FUMBLE", "DOWNS")) %>%
  filter(series_result %in% c("Touchdown", "Field goal")) %>% 
  select(posteam, defteam, drive_start_transition, desc, series_result) %>% 
  group_by(posteam) %>% 
  summarize(total=n())
```

```{r}
scoresofftovs10 <- nfl10 %>% 
  filter(str_detect(desc, "is GOOD")) %>% 
  filter(drive_start_transition %in% c("INTERCEPTION", "FUMBLE", "DOWNS")) %>%
  filter(series_result %in% c("Touchdown", "Field goal")) %>% 
  select(posteam, defteam, drive_start_transition, desc, series_result) %>% 
  group_by(posteam) %>% 
  summarize(total=n())
```

```{r}
scoresofftovs09 <- nfl09 %>% 
  filter(str_detect(desc, "is GOOD")) %>% 
  filter(drive_start_transition %in% c("INTERCEPTION", "FUMBLE", "DOWNS")) %>%
  filter(series_result %in% c("Touchdown", "Field goal")) %>% 
  select(posteam, defteam, drive_start_transition, desc, series_result) %>% 
  group_by(posteam) %>% 
  summarize(total=n())
```

```{r}
scoresofftovs08 <- nfl08 %>% 
  filter(str_detect(desc, "is GOOD")) %>% 
  filter(drive_start_transition %in% c("INTERCEPTION", "FUMBLE", "DOWNS")) %>%
  filter(series_result %in% c("Touchdown", "Field goal")) %>% 
  select(posteam, defteam, drive_start_transition, desc, series_result) %>% 
  group_by(posteam) %>% 
  summarize(total=n())
```

```{r}
scoresofftovs07 <- nfl07 %>% 
  filter(str_detect(desc, "is GOOD")) %>% 
  filter(drive_start_transition %in% c("INTERCEPTION", "FUMBLE", "DOWNS")) %>%
  filter(series_result %in% c("Touchdown", "Field goal")) %>% 
  select(posteam, defteam, drive_start_transition, desc, series_result) %>% 
  group_by(posteam) %>% 
  summarize(total=n())
```

```{r}
scoresofftovs06 <- nfl06 %>% 
  filter(str_detect(desc, "is GOOD")) %>% 
  filter(drive_start_transition %in% c("INTERCEPTION", "FUMBLE", "DOWNS")) %>%
  filter(series_result %in% c("Touchdown", "Field goal")) %>% 
  select(posteam, defteam, drive_start_transition, desc, series_result) %>% 
  group_by(posteam) %>% 
  summarize(total=n())
```

```{r}
scoresofftovs05 <- nfl05 %>% 
  filter(str_detect(desc, "is GOOD")) %>% 
  filter(drive_start_transition %in% c("INTERCEPTION", "FUMBLE", "DOWNS")) %>%
  filter(series_result %in% c("Touchdown", "Field goal")) %>% 
  select(posteam, defteam, drive_start_transition, desc, series_result) %>% 
  group_by(posteam) %>% 
  summarize(total=n())
```

```{r}
scoresofftovs04 <- nfl04 %>% 
  filter(str_detect(desc, "is GOOD")) %>% 
  filter(drive_start_transition %in% c("INTERCEPTION", "FUMBLE", "DOWNS")) %>%
  filter(series_result %in% c("Touchdown", "Field goal")) %>% 
  select(posteam, defteam, drive_start_transition, desc, series_result) %>% 
  group_by(posteam) %>% 
  summarize(total=n())
```

```{r}
scoresofftovs03 <- nfl03 %>% 
  filter(str_detect(desc, "is GOOD")) %>% 
  filter(drive_start_transition %in% c("INTERCEPTION", "FUMBLE", "DOWNS")) %>%
  filter(series_result %in% c("Touchdown", "Field goal")) %>% 
  select(posteam, defteam, drive_start_transition, desc, series_result) %>% 
  group_by(posteam) %>% 
  summarize(total=n())
```

```{r}
scoresofftovs02 <- nfl02 %>% 
  filter(str_detect(desc, "is GOOD")) %>% 
  filter(drive_start_transition %in% c("INTERCEPTION", "FUMBLE", "DOWNS")) %>%
  filter(series_result %in% c("Touchdown", "Field goal")) %>% 
  select(posteam, defteam, drive_start_transition, desc, series_result) %>% 
  group_by(posteam) %>% 
  summarize(total=n())
```

```{r}
scoresofftovs <- rbind (scoresofftovs02, scoresofftovs03, scoresofftovs04, scoresofftovs05, scoresofftovs06, scoresofftovs07, scoresofftovs08, scoresofftovs09, scoresofftovs10, scoresofftovs11, scoresofftovs12, scoresofftovs13, scoresofftovs14, scoresofftovs15, scoresofftovs16, scoresofftovs17, scoresofftovs18, scoresofftovs19, scoresofftovs20, scoresofftovs21) %>% 
  group_by(posteam) %>% 
  summarize(sum(total))
```

```{r}
bigdf <- teamtakeawaysince02 %>% inner_join(scoresofftovs, by=c('team'))
```

```{r}
bigdf <- bigdf %>% 
  rename(takeaways='sum(takeaways)', scores_off_takeaways='sum(total)')
```

```{r}
bigdf <- bigdf %>% 
  mutate(pctscoresofftakeaways=(scores_off_takeaways/takeaways)*100) %>% 
  arrange(desc(pctscoresofftakeaways))
```

## Based on what I calculated, from 2002-2021, the Ravens scored most frequently after forcing a turnover (42.7%), while the Browns scored the least frequently after doing so (33.95%). Since 2002, the Patriots have forced the most turnovers (836) while the Raiders have forced the fewest (517). Additionally, since 2002 the Patriots have scored the most times (345) after forcing a turnover, while the Lions have scored the fewest times (196). ##