---
title: "RealFinalCBBPresentation"
author: "Tanner Malinowski"
date: "2022-12-08"
output: html_document
---

On April 4, 2011, UConn basketball shot 19-of-55 (34.5%) from the field and made just one 3-pointer on 11 attempts, a putrid 9.1% clip from beyond the arc. The Huskies finished with 11 turnovers against six assists and scored just 53 points. But when the final buzzer sounded that night, UConn had secured college basketball's crown jewel: a national championship. It wasn't pretty, but UConn's offensive output was more than sufficient against a Butler team that somehow nearly shot twice as worse from the field (12-64, a frigid 18.8%) and managed just 41 points. Since Butler had come oh-so-close to a National Championship the year before, I was pulling for Brad Stevens' squad to avenge last season's (literal) near-miss. Throughout the 2010-11 season, the Bulldogs averaged 71.2 points per game and had scored fewer than 53 points just twice, with their season-low being 49 points. In the NCAA Tournament leading up to the National Championship, the Bulldogs averaged 67.2 points per game and hadn't scored fewer than 60 points. It hurt knowing that Butler needed just 54 points -- 13 points below their tournament average and 17 below their season average -- to erase the sting of last season's loss to Duke. Of course, UConn's defense deserves plenty of credit for stifling Butler. The Huskies' 10 blocks remain a national championship game record. But the Bulldogs only turned it over six times and shot just 8-14 from the line. Their 52 misses from the field were by far the most they logged in a game that season. Not to take anything away from UConn, but there's a tiny part of me that feels like they didn't fully deserve it. Of course, their defense was terrific. But the Bulldogs left six points at the line and chucked up a whopping 42 shots that weren't blocked, but missed nonetheless. Overall, I wanted to know how rare it's been for a college basketball team to win despite shooting as poorly as UConn did. Also, I wanted to know how teams who shoot so poorly manage to win the game. Is it because of a stifling defense/even worse offensive showing by the other team, or something else?

```{r}
library(tidyverse)
library(ggplot2)
library(ggbeeswarm)
library(ggrepel)
library(waffle)
```

This dataframe loaded information from every men's college basektball game from the 2011-12 season through the 2021-22 season. 
```{r}
basketball <- read_csv("https://dwillis.github.io/sports-data-files/logs1222.csv")
```

I noticed that the dataframe included each game for every team, meaning each game should be listed twice: once for the winning team and once for the losing team. But I also noticed that there were an odd number of rows, meaning not every game was accounted for twice.

I dove into the win-loss column to see if I could find anything explaining it.
```{r}
basketball %>% 
  group_by(W_L) %>% 
  summarize(total=n())
```
Already, I noticed there were 22 results that were supposedly undecided.

```{r}
basketball %>% 
  filter(is.na(W_L))
```
There were 11 games from the 2021-22 season that were basically unaccounted for.

```{r}
basketball_cleaned <- basketball %>% 
  filter(str_detect(W_L, "W") | str_detect(W_L, "L"))
```
Still, that took away 22 rows, which left me with another odd number.

```{r}
basketball_cleaned %>% 
  group_by(W_L) %>% 
  summarize(total=n())
```
From here I noticed that the W_L column had 4000 more "W's" than "L's." I figured this was because the dataframe accounted for a Division 1 school's win over a Division 3 team, but not the other way around.

```{r}
basketball_cleaned %>% 
  summarize(n_distinct(TeamFull))
```


```{r}
basketball_cleaned %>% 
  summarize(n_distinct(Opponent))
```
That was probably the case, as there were 813 more unique Opponents than TeamFulls.

I began trying even out the columns to get the same number of distinct teams, but I realized it was too much of a hassle. 
```{r}
basketball_cleaned %>% 
  group_by(TeamFull) %>% 
  summarize(total=n()) %>% 
  arrange(total)
```

```{r}
basketball_cleaned %>% 
  group_by(Opponent) %>%
  filter(n() > 238) %>% 
  summarize(total=n()) %>% 
  arrange(total)
```

I made things easier on myself by just including games that had been won by somebody.
```{r}
basketball_winners <- basketball_cleaned %>% 
  filter(str_detect(W_L, "W")) %>% 
  rename(winner = TeamFull, loser = Opponent)
```

So, throughout the 2011-12 and 2021-22 NCAA men's basketball seasons, there had been 63,352 games won. 

Games where both teams shot below 34.6% from the floor: 562
```{r}
basketball_winners %>% 
  filter(TeamFGPCT < 0.346, OpponentFGPCT < 0.346)
```

Games where just the winning team shot below 34.6% from the floor: 808
```{r}
basketball_winners %>% 
  filter(TeamFGPCT < 0.346, OpponentFGPCT > 0.345)
```

Games where just the losing team shot below 34.6% from the floor: 12,413
```{r}
basketball_winners %>% 
  filter(TeamFGPCT > 0.345, OpponentFGPCT < 0.346)
```


Total games where a team shot below 34.6% from the floor: 13,783

Overall, there were 13,783 games from the 2011-12 season through the 2021-22 season in which a team shot below 34.6% from the floor. In other words, 21.8% of games during that time saw at least one team shoot worse than 34.6% from the floor.
```{r}
field_goals <- c("BothAbove"=49569, "LoserBelow"=12413, "WinnerBelow"=808, "BothBelow"=562)
```

```{r}
waffle(
      field_goals/150, 
      rows = 15, 
      title="It Was Rare for a Team to Shoot Below 34.6% From the Field",
      xlab="1 square = 150 games", 
      colors = c("black", "red", "grey", "blue")
)
```


```{r}
```{r}
below34.6pct <- basketball_winners %>% 
  filter(
    TeamFGPCT < 0.346 | OpponentFGPCT < 0.346
  )

winsbelow34.6pct <- below34.6pct %>% 
  filter(str_detect(W_L, "W"), TeamFGPCT <0.346)
```

```{r}
teamshotbelow34.6pct <- c("ShotBelow34.6%" = 13783, "ShotAbove34.5%" = 49569)
teamshotbelow34.6pctandwon <- c("ShotBelow34.6%andWon" = 1370, "ShotBelow34.6%andLost" = 12413)
```

```{r}
waffle(
      teamshotbelow34.6pct/100, 
      rows = 22, 
      title="Teams Shot 34.5% or below in 21.8% of games",
      xlab="1 square = 100 games", 
      colors = c("black", "red")
)

waffle(
      teamshotbelow34.6pctandwon/25, 
      rows = 22, 
      title="Teams that Shot 34.5% or Below Rarely Won",
      xlab="1 square = 25 games", 
      colors = c("black", "red")
)
```

What about 3-point percentage? UConn shot just 9.1% from deep. 
```{r}
basketball_winners %>% 
  filter(Team3PPCT < 0.092 | Opponent3PPCT < 0.092)
```
A team shot 9.1% or worse from 3-point range just 1,684 times throughout 63,352 games.

Games where both teams shot below 9.2% from three: 12
```{r}
basketball_winners %>% 
  filter(Team3PPCT < 0.092, Opponent3PPCT < 0.092)
```

Games where just the winning team shot below 9.2% from three: 381
```{r}
basketball_winners %>% 
  filter(Team3PPCT < 0.092, Opponent3PPCT > 0.092)
```

Games where just the losing team shot below 9.2% from three: 1291
```{r}
basketball_winners %>% 
  filter(Team3PPCT > 0.092, Opponent3PPCT < 0.092)
```

Total games where a team shot below 9.2% from three: 1684
```{r}
basketball_winners %>% 
  filter(Team3PPCT < 0.092 | Opponent3PPCT < 0.092)
```

```{r}
threes <- c("BothAbove"=61668, "LoserBelow"=1291, "WinnerBelow"=381, "BothBelow"=12)
```

I broke this down to show all four distinctions throghout the 2011-12 and 2021-22 seasons.
```{r}
waffle(
      threes/150, 
      rows = 15, 
      title="It Was Rare for a Team to Shoot Below 9.2% from 3-point Range",
      xlab="1 square = 150 games", 
      colors = c("black", "red", "grey", "blue")
)
```

```{r}
shot9.1pctorworse <- c("Shot9.1pctOrWorse"=1684, "Shot9.2pctOrBetter"=61668)
shot9.1pctorworseandwon <- c("Shot9.1pctorWorseAndWon" = 393, "Shot9.1pctorWorseAndLost"=1291)
```

I broke this down to show 1) how often teams shot below 9.2% from 3-point range from 2011-12 - 2021-22, and 2) how frequently teams won despite shooting below 9.2% from three.
```{r}
waffle(
      shot9.1pctorworse/250, 
      rows = 14, 
      title="Teams Shot 9.1% or Below on 3-pointers in 2.7% of games",
      xlab="1 square = 250 games", 
      colors = c("black", "red")
)

waffle(
      shot9.1pctorworseandwon/5, 
      rows = 14, 
      title="Teams that Shot < 9.2% on 3-pointers won 23.3% of the time",
      xlab="1 square = 5 games", 
      colors = c("black", "red")
)

```

```{r}
field_goals <- c("BothAbove"=49569, "LoserBelow"=12413, "WinnerBelow"=808, "BothBelow"=562)
```

```{r}
waffle(
      field_goals/150, 
      rows = 15, 
      title="It Was Rare for a Team to Shoot Below 34.6%",
      xlab="1 square = 100 games", 
      colors = c("black", "red", "grey", "blue")
)
```










```{r}
How did UConn fare compared to other national championship game winners?
```{r}
champs <- basketball %>% 
  filter(
    W_L == "W",
    Date == "2012-04-02" |
    Date == "2013-04-08" |
    Date == "2014-04-07" |
    Date == "2015-04-06" |
    Date == "2016-04-04" |
    Date == "2017-04-03" |
    Date == "2018-04-02" |
    Date == "2019-04-08" |
    Date == "2021-04-05" |
    Date == "2022-04-04"
  )
```


```{r}
ggplot() + 
  geom_point(
    data=champs, 
    aes(x=TeamFGPCT, y=Team3PPCT, size=TeamScore), 
    color="red", 
    alpha=.5) + 
   geom_text_repel(
    data=champs, 
    aes(x=TeamFGPCT, y=Team3PPCT, label=TeamFull)
  )+
  labs(title="What pages did UNC steal out of UConn's playbook?", subtitle="UNC did not help itself much on offense, but it didn't matter in the end", caption="Source: Derek Willis | By Tanner Malinowski")  + theme_minimal() + 
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 8), 
    plot.subtitle = element_text(size=10), 
    panel.grid.minor = element_blank()
    )
```
```{r}
bad_three_point_winners <- bad_three_point_games %>%
  filter(Team3PPCT < 0.092)
```

```{r}
bad_three_point_winners %>% 
  summarize(avg=mean(Team3PPCT))
```
```{r}
winning_team_three_point_averages <- bad_three_point_winners %>% 
  summarize(avg=mean(Team3PPCT))
```

```{r}
losing_team_three_point_averages <- bad_three_point_winners %>% 
  summarize(avg=mean(Opponent3PPCT))
```

```{r}
three_point_average_bar <- full_join(winning_team_three_point_averages, losing_team_three_point_averages, by = "avg")
```


```{r}
ggplot() + 
  geom_bar(
    data=three_point_average_bar, 
    aes(x=avg,
        weight = `avg`))



ggplot() + 
  geom_bar(
    data=three_point_average_bar, 
    aes(x=reorder(avg, -`avg`),
        weight=`avg`)) + 
  scale_y_continuous(labels=comma) + 
  labs(
    title="3-Point Percentage", 
    x="Teams", 
    y="Three Point Percentage") + 
  theme_minimal()
```

```{r}

```















What about UConn's 3-point shooting?
```{r}
bad_three_point_games <- basketball_winners %>% 
  filter(Team3PPCT < 0.092 | Opponent3PPCT < 0.092)
```

1,684 out of 63,352 games saw at least one team shoot below 9.2% from 3-point range.

```{r}
wins_below_34.6_fg_pct_and_below_9.2_3pt_pct <- rbind(winsbelow34.6pct, bad_three_point_winners)
```

```{r}
bad_fgs_and_threes <- wins_below_34.6_fg_pct_and_below_9.2_3pt_pct %>% 
  filter(TeamFGPCT <0.346, Team3PPCT <0.092)
```
There were 90 games throughout the 2011-12 season and 2021-22 season where a team won despite shooting below 34.6% from the field and below 9.2% from 3-point range.

```{r}
bad_fgs_and_threes %>% 
  mutate(ast_to_tov_differential = TeamAssists - TeamTurnovers) %>% 
  arrange(ast_to_tov_differential)
```


```{r}
ggplot() + 
  geom_bar(
    data=bad_fgs_and_threes, 
    aes(
      x=reorder(Winner, -`2019`), 
      weight=`2019`)
    ) + 
  scale_y_continuous(labels=comma) + 
  labs(
    title="Top 10 Football Programs By Attendance", 
    x="School", 
    y="Attendance"
)
```
























































































simple_below34.6pct <- below34.6pct %>% 
  group_by(Season) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total)) %>% 
  mutate(type="GamesShootingBelow34.6")
```

```{r}
simple_winsbelow34.6pct <- winsbelow34.6pct %>% 
  group_by(Season) %>% 
  summarize(total=n()) %>% 
  arrange(desc(total)) %>% 
  mutate(type="WinsShootingBelow34.6")
```

```{r}
basketball_games <- basketball_winners %>% 
    group_by(Season) %>% 
    summarize(total=n()) %>% 
    arrange(desc(total)) %>% 
    mutate(type="TotalGames")
```

```{r}
games_and_bad_shooting <- rbind(basketball_games, simple_below34.6pct) %>% 
  group_by(Season)

games_and_wins_below_34.6 <- rbind(simple_below34.6pct, simple_winsbelow34.6pct) %>% 
  group_by(Season)




